Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-10-27 forecast/paper/paper.tex
Finish minor paper mods; all but SS, abstract. Code for hub comparison

While forecasting carries a rich literature that offers a wide range of
techniques, see, e.g., \cite{Hyndman:2018},
% put Ioannidis ref here
we purposely constrain ourselves to 
very simple models, avoiding common enhancements such as order selection,
correction of outliers/anomalies in the data, and inclusion of regularization or
nonlinearities. For the same reason, we do not account for other factors that
may well aid in forecasting, such as age-specific effects, holiday adjustments,
public health mandates, and so on. That said, analyses of forecasts submitted
to the COVID-19 Forecast Hub \cite{ForecastHub} by a large community of modelers 
have shown that simple, robust models have consistently been among the
best-performing over the pandemic, including time series models similar the ones
we use in this paper. See, e.g., \cite{Cramer:2021}, for a detailed study
of death forecasting. 
% added
Similarly, the simple autoregressive models used here for case forecasting
are competitive with top models submitted to the Forecast Hub over the same
period of time (see the Supplement for this analysis).
% Our internal evaluations suggest that the same is true (at
% least, if not to a greater extent) for case forecasting.

@@

There is another, more subtle issue in evaluating predictive utility that 
deserves explicit mention, as it will play a key role in our analysis.
Signals computed from surveillance streams will often be subject to  
latency and/or revision. For example, a signal based on aggregated medical
insurance claims may be available after just a few days, but it can then be
substantially revised over the next several weeks as additional claims are
submitted and/or processed late. Correlations between such a signal and case
rates calculated ``after the fact'' (i.e., computed retrospectively, using the
finalized values of this signal) will not deliver an honest answer to the    
question of whether this signal would have been useful in real time. Instead,
we build predictive models using only the data that would have been available
\textit{as of} the prediction date, and compare the ensuing predictions in terms 
of accuracy. To do so, we leverage Delphi's \texttt{evalcast} R package 
\cite{EvalcastR}, which plugs into the COVIDcast API's data versioning system,  
and facilitates honest backtesting.
%
TODO: Here we cite many refs
\cite{faust2009a,diebold1991a,patterson1995a,croushore2011a,croushore2006a,harvey1983a,mankiw1986a,trivellato1986a,clark2009a,koenig2003a,howrey1978a}

@@

In the supplement, we show how to define a quantitative score to measure the
leadingness of an indicator, at any time $t$ and any location $\ell$, based on
cross correlations to case rates over a short time window around $t$.
The higher this score, the greater it ``leads'' case activity.
%
TODO: Here we cite many refs.
\cite{yeats1972a,emerson1996a,koch1988a,sargent1977a,engle1987a,shiskin1968a,granger1969a,lahiri1991a,koopmans1947a,StockWatson1989,diebold1989a,mitchell1938a,auerbach1982a,hamilton1996a}
%
Figure 
\ref{fig:leading} displays correlations between the leadingness score of an 
indicator and the WIS difference (AR model minus an indicator model), stratified
by whether the target is classified as up, down, or flat.  One would naturally
expect that the WIS difference would be positively correlated with leadingness.
Somewhat surprisingly, this relationship turns out to be strongest in down
periods and weakest in up periods.  In fact, it is very nearly the case that for
each indicator, the strength of correlations only decreases as we move from down
to flat to up  periods.  In the supplement, we extend this analysis by studying 
analogous ``laggingness'' scores, but we do not find as clear patterns.

>>>>

While forecasting carries a rich literature that offers a wide range of
techniques, see, e.g., \cite{Hyndman:2018},
we purposely constrain ourselves to 
very simple models, avoiding common enhancements such as order selection,
correction of outliers/anomalies in the data, and inclusion of regularization or
nonlinearities. For the same reason, we do not account for other factors that
may well aid in forecasting, such as age-specific effects, holiday adjustments,
public health mandates, and so on.
% Paragraph from Rob
Work in \cite{ioannidis2020a} points out many additional difficulties when it
comes to
forecasting COVID-19 case rates and deaths. These include bad data, wrong
assumptions, lack of incorporation of epidemiological features, etc.  We agree
with most of these 
and our group has found that COVID-19 forecasting to be extremely challenging.
However the point of our  paper is to demonstrate the utility of  auxiliary
indicators for improving COVID-19 forecasting, and to suggest that these
indicators might be useful for future pandemics. We are not suggesting that
these indicators make COVID-19 forecasting  easy. 
Despite these obstacles, analyses of forecasts submitted
%
to the COVID-19 Forecast Hub \cite{ForecastHub} by a large community of modelers 
have shown that simple, robust models have consistently been among the
best-performing over the pandemic, including time series models similar the ones
we use in this paper. See, e.g., \cite{Cramer:2021}, for a detailed study
of death forecasting. 
% added
Similarly, the simple autoregressive models used here for case forecasting
are competitive with top models submitted to the Forecast Hub over the same
period of time (see the Supplement for this analysis).
% Our internal evaluations suggest that the same is true (at
% least, if not to a greater extent) for case forecasting.

@@

There is another, more subtle issue in evaluating predictive utility that 
deserves explicit mention, as it will play a key role in our analysis.
Signals computed from surveillance streams will often be subject to  
latency and/or revision. For example, a signal based on aggregated medical
insurance claims may be available after just a few days, but it can then be
substantially revised over the next several weeks as additional claims are
submitted and/or processed late. Correlations between such a signal and case
rates calculated ``after the fact'' (i.e., computed retrospectively, using the
finalized values of this signal) will not deliver an honest answer to the    
question of whether this signal would have been useful in real time. Instead,
we build predictive models using only the data that would have been available
\textit{as of} the prediction date, and compare the ensuing predictions in terms 
of accuracy. To do so, we leverage Delphi's \texttt{evalcast} R package 
\cite{EvalcastR}, which plugs into the COVIDcast API's data versioning system,  
and facilitates honest backtesting.
%
The necessity of real-time data for honest forecast evaluations has a long history in econometrics,
\cite[see,
e.g.,][]{faust2009a,diebold1991a,patterson1995a,croushore2011a,croushore2006a,harvey1983a,mankiw1986a,trivellato1986a,clark2009a,koenig2003a,howrey1978a}
but it is far less acknowledged in public health despite its demonstrable importance.

@@

In the supplement, we define a quantitative score to measure the
leadingness of an indicator, at any time $t$ and any location $\ell$, based on
cross correlations to case rates over a short time window around $t$.
The higher this score, the greater it ``leads'' case activity.
%
This analysis is closely related to Granger Causality \cite{engle1987a} and
draws on a large body of prior work that measures leadingness in economic time-series 
\cite[e.g.,][]{yeats1972a,emerson1996a,koch1988a,sargent1977a,engle1987a,shiskin1968a,granger1969a,lahiri1991a,koopmans1947a,StockWatson1989,diebold1989a,mitchell1938a,auerbach1982a,hamilton1996a}.
%
Figure 
\ref{fig:leading} displays correlations between the leadingness score of an 
indicator and the WIS difference (AR model minus an indicator model), stratified
by whether the target is classified as up, down, or flat.  One would naturally
expect that the WIS difference would be positively correlated with leadingness.
Somewhat surprisingly, this relationship turns out to be strongest in down
periods and weakest in up periods.  In fact, it is very nearly the case that for
each indicator, the strength of correlations only decreases as we move from down
to flat to up  periods.  In the supplement, we extend this analysis by studying 
analogous ``laggingness'' scores, but we do not find as clear patterns.
